---
title: "R Notebook"
output: html_notebook
---

# now will take inspiration from Joanna and get the tn5 insertion sites and save the coverage as a bigwig

# but first have to check if I can skip the steps for bam files that already had their reads shifted to account for the tn5 gap

# essentially this script should be for bam files that already used a tool like deeptools alignmentsieve --ATACshift

# if I want to use this script in nextflow I would just make a channel that has all the bams and input it in a process that runs the r code natively and not a script

```{r}
library(GenomicAlignments)
library(GenomicRanges)

library(rtracklayer)


# load the atac shifted bams
getwd()
atac_shifted_bams = list.files(path = "./test_atac_data_pipeline/atac_bams/atac_shifted_bams/", pattern = "*ATAC.*bam$", full.names = TRUE)

list_of_names = c()
for (i in atac_shifted_bams){
  # print(i)
  names_of_files = unlist(str_split(basename(i), pattern = "\\."))[1]
  # print(names_of_files)
  
  list_of_names = append(list_of_names, names_of_files) 
  print(list_of_names)
  
}


#import the size factor file
atac_seq_sizefactor_table = read.table(file = "./test_atac_data_pipeline/nextflow_R_script_outputs/ATAC/size_factor_calculation.tsv", sep = "\t", header = TRUE)

names(atac_shifted_bams) = list_of_names
# now I can take the bam files and use genomic alignment to load in the reads and resize them to only the start coordinate
num = 1 
# list_of_readcuts_coverage = c()
for (atac_bam in atac_shifted_bams){
  
  print(names(atac_shifted_bams[num]))
  bam_data = readGAlignmentPairs(atac_bam)
  
  # getting the forward reads
  forward_reads = bam_data@first
  
  # getting the reverse reads
  reverse_reads = bam_data@last
  
  # resize to only get the start positions of the forward and reverse reads, but use fix="start" so it is strand aware and thus gets the end position as the start position for the reverse reads.
  forward_read_start = resize(GRanges(forward_reads), fix = "start", width = 1)
  
  reverse_read_start = resize(GRanges(reverse_reads), fix = "start", width = 1)
  
  # save the reads to a list
  reads_cut = c(forward_read_start, reverse_read_start)
  
  reads_coverage = coverage(reads_cut)
  
  # list_of_readcuts_coverage[[num]] = reads_coverage
  
  
  # lets just make the bigwig here
  # I should normalize by size factor using the sizefactor file I have generated for hugo
  #test_names_sizefactor[grep(paste0(names(atac_shifted_bams)[2],"*"), test_names_sizefactor)]
  # this above is something that works where i get the name from the current loop, and paste a star at the end to grep the correct index from the size factor table
  
  # here I get the name and see what index it is in the size factor table
  correct_file_index = grep(paste0(names(atac_shifted_bams)[num],"*"), atac_seq_sizefactor_table[,1]) 
  
  # then take that correct file index and get the size factor number from the 2 column in that row
  correct_sizeFactor_num = atac_seq_sizefactor_table[correct_file_index, 2]
  
  # now multiply the coverage by this size factor
  normalized_reads_coverage = reads_coverage * correct_sizeFactor_num
  
  rtracklayer::export.bw(reads_coverage, con = paste0("./test_atac_data_pipeline/",names(atac_shifted_bams)[num], "_CutsCoverage.bigwig"))
  
  # the normalized bigwigs
  rtracklayer::export.bw(normalized_reads_coverage, con = paste0("./test_atac_data_pipeline/", "normalized_by_SF_", names(atac_shifted_bams)[num], "_CutsCoverage.bigwig"))
  
  
  
  num = num+1
}


```

# see if there is a deeptools library for plotting in R

```{r}




```

