---
title: "using IDR peaks"
output: html_notebook
---



```{r}

hlow_idr_peaks = list.files(path="./idr_results/H3k27me3/H1low/", pattern = "IDR.*broadPeak$", full.names = TRUE )
print(hlow_idr_peaks)

# now find which broadPeak file has the most lines ( the most amount of peaks called)

hlow_peak_counts = data.frame(file = hlow_idr_peaks,
                            peak_counts = sapply(hlow_idr_peaks, function(x) {length(readLines(x))} )
                            
                            )

hlow_peak_counts

best_hlow_idr = hlow_peak_counts[which.max(hlow_peak_counts$peak_counts),]$file

# now this gives the file with the most peaks as according to encode pipeline
best_hlow_idr


# now doing the same for scrm

scrm_idr_peaks = list.files(path="./idr_results/H3k27me3/Scrm/", pattern = "IDR.*broadPeak$", full.names = TRUE )
print(scrm_idr_peaks)

# now find which broadPeak file has the most lines ( the most amount of peaks called)

scrm_peak_counts = data.frame(file = scrm_idr_peaks,
                            peak_counts = sapply(scrm_idr_peaks, function(x) {length(readLines(x))} )
                            
                            )

scrm_peak_counts

best_scrm_idr = scrm_peak_counts[which.max(scrm_peak_counts$peak_counts),]$file
```

## Now do I use all of the bam files to get counts with the chromVAR get counts package or do I only use the 4 libraries that went into making the idr pair peak files?

# this is me getting the correct bam files 
```{r}

# now lets use chromVAR to load in the counts

# load the bam files separately per condition so i can choose the correct one

hlow_bam_files = list.files(path="./atac_bams/", pattern="dH1.*ATAC.*bam$", full.names = TRUE )
hlow_bam_files

# now get the tokens from the best hlow file and see if it is in the name of the bam. if it is, we keep the bam

hlow_idr_tokens = strsplit(best_hlow_idr,split = "_")[[1]]
first_rep_hlow = hlow_idr_tokens[5]
second_rep_hlow = hlow_idr_tokens[7]


# make an empty list to store the master bams
list_of_mBams = list()

for (x in hlow_bam_files) {
  
  target_bam = strsplit(x, split = "_")[[1]]
  
  if (first_rep_hlow %in% target_bam) {
    
    print(paste("first rep is here", x))
    
    list_of_mBams = append(list_of_mBams, x)
  }
  if (second_rep_hlow %in% target_bam) {
    
    print(paste("second rep is here", x))
    
    list_of_mBams = append(list_of_mBams, x)
    
  }
  
}

list_of_mBams

# now doing the same for scrm


scrm_bam_files = list.files(path="./atac_bams/", pattern="scr.*ATAC.*bam$", full.names = TRUE )
scrm_bam_files

# getting tokens
scrm_idr_tokens = strsplit(best_scrm_idr,split = "_")[[1]]
first_rep_scrm = scrm_idr_tokens[5]
second_rep_scrm = scrm_idr_tokens[7]


for (x in scrm_bam_files) {
  
  target_bam = strsplit(x, split = "_")[[1]]
  
  if (first_rep_scrm %in% target_bam) {
    
    print(paste("first rep is here", x))
    
    list_of_mBams = append(list_of_mBams, x)
  }
  if (second_rep_scrm %in% target_bam) {
    
    print(paste("second rep is here", x))
    
    list_of_mBams = append(list_of_mBams, x)
    
  }
  
}



```




# start from here with the concat 0.4 idr peaks since you dont have to find the peak file with the most peaks anymore

```{r}
# now that i have the hlow and scrm that have the most peaks called from the idr pairs
# I will create the master peak file

#peaklist = list(best_hlow_idr, best_scrm_idr )
#peaklist = list('./idr_results/H3k27me3/H1low/concat_IDR_H1low_H3k27me3_r1_vs_r2_vs_r3_0.4_pairs.broadPeak', './idr_results/H3k27me3/Scrm/concat_IDR_Scrm_H3k27me3_r1_vs_r2_vs_r3_0.4_pairs.broadPeak')

#peaklist = list('./idr_results/H3k27me3/H1low/concat_IDR_H1low_H3k27me3_r1_vs_r2_vs_r3_0.4_pairs.broadPeak', './idr_results/H3k27me3/published/concat_IDR_published_H3k27me3_r1_vs_r2_vs_null_0.4_pairs.broadPeak')

#best_hlow_idr
#best_scrm_idr

library(GenomicRanges)
library(chromVAR)
library(DESeq2)
library(tidyr)
library(EnhancedVolcano)

# making the master peak genomic ranges object
# mPeak = GRanges()
# 
# for (peakfile in peaklist) {
# 
#   peaktable = read.table(peakfile, header = FALSE, sep = "\t")
# 
#   gr_object = GRanges(seqnames = peaktable$V1, IRanges(start = peaktable$V2, end = peaktable$V3), strand = "*" )
# 
#   mPeak = append(mPeak, gr_object)
# }
# 
# # making sure there are no redundant peaks
# masterPeak = reduce(mPeak)

# might have to let it auto detect the file type when using narrow peaks that arent merged into a strict bed file
mPeak = rtracklayer::import('./idr_results/merged_concat_peaks/concat_master_bothConditions_ATAC_ATAC.bed', format = "BED")

# using resize to extnd regions
extended_master_peak = resize(mPeak, width = width(mPeak)+1, fix = "center")

# before trimming i need to add the sequence lengths of the correct genome
# will have to find a way to automate this step for other genomes


masterPeak_beta = reduce(extended_master_peak)
# now to resize again to remove the artificial 50kb from the start and end of the newly reduced peak

new_peak_size = width(masterPeak_beta)-1
masterPeak = resize(masterPeak_beta,  width = new_peak_size , fix = "center")


```


# doing this for the concat files just use all bams

```{r}

hlow_bam_files = list.files(path="./atac_bams/", pattern="dH1.*ATAC.*bam$", full.names = TRUE )
hlow_bam_files

scrm_bam_files = list.files(path="./atac_bams/", pattern="scr.*ATAC.*bam$", full.names = TRUE )
scrm_bam_files

list_of_mBams = c(hlow_bam_files, scrm_bam_files)

list_of_mBams
```


## Now to use chromVAR to fill the matrix


```{r}

# making the matrix 

countsMatrix = matrix(data = NA, length(masterPeak), length(list_of_mBams) )


# I want to get the chr start end of the peaks and put them in the matrix as column names so I know which peaks I am looking at from the deseq2 results in the differential peaks.
seqnames(masterPeak)

# i should put the unique ids from the master peak object in the matrix as row names.
unique_masterPeak_ids = paste0(as.character(seqnames(masterPeak)), ":" ,start(masterPeak), "-", end(masterPeak))

rownames(countsMatrix) = unique_masterPeak_ids

#getting the list of bam base names to add to the matrix column names
list_bam_basenames = list()

library(Rsubread)

# featureCounts in rsubread needs a dataframe and in the correct format, with a GeneID column first then chr1 column
df_masterPeak = as.data.frame(masterPeak)

# now I need to change the second column name to chr
df_masterPeak = df_masterPeak %>% rename(chr = seqnames)

# next i need to add the unique ids as the columns in this dataframe
df_masterPeak$GeneID = unique_masterPeak_ids


# for deseq2 i need the condition design. so hlow and scrm
# then i will find a way to tally how many of each are there so i can automate the rep count
condition_design = list()

type_design = list()

for (x in c(1:length(list_of_mBams))) {
  
  path_bam = list_of_mBams[[x]]
  print(path_bam)
  bam_basename = basename(path_bam)
  bam_tokens = strsplit(basename(path_bam), split = "_")[[1]]
  
  # labeling the important tokens so it is easier to keep track of
  condition = bam_tokens[1]
  histone = bam_tokens[2]
  replicate = bam_tokens[3]
  
  
  # for later parts I need the condition and to know how many times to repeat them
  condition_design = append(condition_design, paste(condition,histone,sep="_"))
  
  
   # also get the replicate too for type design
  type_design = append(type_design, replicate)
  #type_design = append(type_design, paste(histone,replicate, sep="_"))
  
  
  
  # using chromVAR getCounts function
  # fragment_counts = getCounts(path_bam, masterPeak, paired= TRUE, by_rg = FALSE, format = "bam")
  
  fragment_counts <- featureCounts(
    files = path_bam,
    annot.ext = df_masterPeak,
    isPairedEnd = TRUE,      # TRUE if your data were paired-end
    strandSpecific = 0,       # 0 = unstranded, 1/2 if strand-specific
    useMetaFeatures = TRUE
  )
  
  
  # putting the fragment counts in the column labeled by the bam name
  countsMatrix[,x] = fragment_counts$counts
  
  list_bam_basenames = append(list_bam_basenames, bam_basename)
  
}

colnames(countsMatrix) = list_bam_basenames  
```

```{r}

#first removing the low count fragments. only keeping the rows that are above 5 count in total

keep_Rows = which(rowSums(countsMatrix) > 5)

filt_countmatrix = countsMatrix[keep_Rows,]

# now to get the condition_design
condition_counts = table(unlist(condition_design))

# this gives back the names and the counts give back the counts for each of the names
condition_names = names(condition_counts)
condition_num = as.numeric(condition_counts)

# now i can put both lists in to get back the experiment design
condition_factor = factor(rep(condition_names, times=condition_num))
# change this if need to switch the experimental design order
condition_factor = relevel(condition_factor, levels(condition_factor)[2])

# repeating the above to have another column with type (replicates)
type_counts = table(unlist(type_design))
type_names = names(type_counts)
type_num = as.numeric(type_counts)

#type_factor = factor(rep(type_names, times=type_counts))
type_factor = factor(rep(type_names, times=type_num[1]))

```

# I want to get the idr threshold and use that as input for the file names and other things

```{r}
#peak_file = basename(peaklist[[1]])
peak_file = './idr_results/merged_concat_peaks/concat_master_bothConditions_ATAC_ATAC.bed'

idr_used = strsplit(peak_file, split = "_")[[1]][10]
idr_used
```

# i think here i can load the up proseq genes and find which ATAC-seq regions are in the TSS of the genes +-5kb
```{r}


proseq_genes = rtracklayer::import('/lustre/fs4/home/rjohnson/pipelines/peak_calling_analysis_pipeline/proseq_data/proseq_up_genes_with_coord.tsv', format = "GFF")

```


# dont use this workflow, use the peaks generated by the pipeline now
# now for deseq2 workflow
```{r}


# now to do the normal deseq2 workflow

dds = DESeqDataSetFromMatrix(countData = filt_countmatrix,
                             colData = DataFrame(condition_factor, type_factor),
                             design = ~ condition_factor)

# using the function on our data
DDS = DESeq(dds)

norm_DDS = counts(DDS, normalized = TRUE) # normalization with respect to the sequencing depth

# adding _norm onto the column names in the normalized matrix
colnames(norm_DDS) = paste0(colnames(norm_DDS), "_norm")


# provides independent filtering using the mean of normalized counts
res = results(DDS, independentFiltering = FALSE, altHypothesis = "greaterAbs")


# this is looking at the differences between the 3 deseq analyzed options
countMatDiff = cbind(filt_countmatrix, norm_DDS, res)

head(countMatDiff)




 # getting the results name and addding to the coef we want to shrink
  experiment_design_name = resultsNames(DDS)[2]
  
  # useful for visualization and ranking of genes or in this case peaks
  resLFC = lfcShrink(DDS, coef= resultsNames(DDS)[2], type = "apeglm")
  
  
  # finding the up and down regulated counts that pass the threshold
  
  all_peaks = resLFC
  total_all_peaks = length(all_peaks[,1])

  up_reg = resLFC[which(resLFC$padj < 0.05 & resLFC$log2FoldChange >= 1) ,]
  total_up_reg = length(up_reg[,1])
  #total_up_reg
  down_reg = resLFC[which(resLFC$padj < 0.05 & resLFC$log2FoldChange <= -1) ,]
  total_down_reg = length(down_reg[,1])
  total_down_reg

  unchanging_reg = resLFC[which(resLFC$padj < 0.05 & resLFC$log2FoldChange <1 & resLFC$log2FoldChange > -1) ,]
  total_unchanging_reg = length(unchanging_reg[,1])
  total_unchanging_reg

  # get the percentage of diff peaks
  up_percentage_reg = round((total_up_reg/total_all_peaks)*100, digits = 1)

  down_percentage_reg = round((total_down_reg/total_all_peaks)*100, 1)

  unchanging_percentage_reg = round((total_unchanging_reg/total_all_peaks)*100, 1)

  # up_reg = resLFC[which(resLFC$padj < 0.05 & resLFC$log2FoldChange >= 1) ,]
  # total_up_reg = length(up_reg[,1])
  # #total_up_reg
  # down_reg = resLFC[which(resLFC$padj < 0.05 & resLFC$log2FoldChange <= -1) ,]
  # total_down_reg = length(down_reg[,1])
  # total_down_reg
  
  resLFC$Label = ifelse(resLFC$padj > 0.05, "not significant", ifelse(abs(resLFC$log2FoldChange) >=1, "|LFC| > 1 & padj < 0.05", "padj < 0.05" ))
  ma_plot_labeled = ggplot(resLFC, aes(x = baseMean, y = log2FoldChange, color=Label))+
    labs(caption = paste("Up reg (red +1 LFC & padj <0.05) = ", total_up_reg, "; percentage: ", up_percentage_reg ,"%" , "\n", "Down reg (red -1 LFC & padj < 0.05) = ", total_down_reg, "; percentage: ", down_percentage_reg, "\nNow the Unchanging total and percentage ", total_unchanging_reg , "%" , "; percentage: " , unchanging_percentage_reg, "%" ), title = experiment_design_name, subtitle = paste("This means the first condition has these points more than in the second condition. The idr threshold for masterPeaks is",idr_used, sep = " "))+
    theme(plot.caption = element_text(size = 25, face = "bold"), axis.text.x = element_text(size = 25), axis.text.y = element_text(size = 25), axis.title.x = element_text(size = 35), axis.title.y = element_text(size = 35), legend.text = element_text(size = 25), title = element_text(size = 25), plot.subtitle = element_text(size = 15))+
    scale_colour_manual(values=c("red", "darkgrey", "pink"))+
    scale_x_continuous(trans='log10')+
    ylim(c(min(-max(resLFC$log2FoldChange),min(resLFC$log2FoldChange)), max(max(resLFC$log2FoldChange),-min(resLFC$log2FoldChange))))+
    geom_point(data = resLFC, aes(x = baseMean, y = log2FoldChange), size = 2)
  
  print(ma_plot_labeled)
  
  
  pdf(file = paste(experiment_design_name,"IDR", idr_used,"MA_plot_all_reps.pdf", sep = "_"), width = 18, height = 10)
  
  print(ma_plot_labeled)
  dev.off()
  
  png(filename = paste(experiment_design_name,"IDR", idr_used,"MA_plot_all_reps.png", sep = "_"), width = 1200, height = 1000, antialias = "subpixel")
  
  print(ma_plot_labeled)
  dev.off()
  
  
  volcano_plot_removed_reps = EnhancedVolcano(resLFC,
                  lab = rownames(resLFC),
                  title = experiment_design_name, subtitle = paste("This means the first condition has these points more than in the second condition. The idr threshold for masterPeaks is",idr_used, sep = " "),
                  x = 'log2FoldChange', FCcutoff = 1,
                  y = 'padj', pCutoff = 0.05, labSize = 3, 
                  
                  )
  
  print(volcano_plot_removed_reps)
  
  png(filename = paste(experiment_design_name,"IDR",idr_used,"volcano_plot_all_reps.pdf", sep = "_"), width = 1200, height = 1000, antialias = "subpixel")
  print(volcano_plot_removed_reps)
  dev.off()
  
  pdf(file = paste(experiment_design_name,"IDR",idr_used,"volcano_plot_all_reps.pdf", sep = "_"), width = 16, height = 10)
  print(volcano_plot_removed_reps)
  dev.off()
  



# using rlog over vst for transformation

rld = rlog(DDS, blind=FALSE)

#it is clear that the rlog transformation inherently accounts for differences in sequencing depth *
head(assay(rld), 5)


library(ggplot2)

pcaData = plotPCA(rld, intgroup=c("condition_factor","type_factor"), returnData = TRUE )
pcaData

percentVar <- round(100 * attr(pcaData, "percentVar"))


pca_plot_rlog = ggplot(pcaData, aes(PC1, PC2, color=condition_factor, shape=type_factor)) +
  ggtitle(paste("PCA plot using Rlog transform IDR", idr_used, sep = " ")) +
  geom_point(size=3) +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
  coord_fixed()
#pca_plot
name_for_rlog_png_file =paste(experiment_design_name,"PCA_plot_IDR", idr_used, "rlog.png", sep = "_")

png(filename = name_for_rlog_png_file, width = 1000, height = 1000, antialias = "subpixel")
print(pca_plot_rlog)
dev.off()

name_for_rlog_pdf_file =paste(experiment_design_name,"PCA_plot_IDR", idr_used, "rlog.pdf", sep = "_")

pdf(file = name_for_rlog_pdf_file, width = 10, height = 10)
print(pca_plot_rlog)
dev.off()


# testing with vst
vsd_t = vst(DDS, blind = FALSE)
head(assay(vsd_t), 5)

pcaData2 = plotPCA(vsd_t, intgroup=c("condition_factor","type_factor"), returnData = TRUE )
pcaData2

percentVar <- round(100 * attr(pcaData2, "percentVar"))
pca_plot_vst = ggplot(pcaData2, aes(PC1, PC2, color=condition_factor, shape=type_factor)) +
  ggtitle(paste("PCA plot using VST transform IDR", idr_used, sep = " ")) +
  geom_point(size=3) +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
  coord_fixed()


#name_for_vst_png_file =paste(experiment_design_name,"PCA_plot_IDR", idr_used, "vst.png", sep = "_")

pdf(file = paste(experiment_design_name,"PCA_plot_IDR",idr_used,"vst_all_reps.pdf", sep = "_"), width = 10, height = 10)
print(pca_plot_vst)
dev.off()

png(file = paste(experiment_design_name,"PCA_plot_IDR", idr_used,"vst_all_reps.png", sep = "_"), width = 1000, height = 1000)
print(pca_plot_vst)
dev.off()

#pca_plot_rlog

#pca_plot_vst


```

# now to do the same chipseeker strat but for ATACseq peaks
```{r}



library(ChIPseeker)

library(TxDb.Hsapiens.UCSC.hg38.knownGene)

txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene

rtracklayer::export.bed(row.names(up_reg), con = "up_ATAC_peaks.bed")

rtracklayer::export.bed(row.names(down_reg), con = "down_ATAC_peaks.bed")

# now exporting the unchanging peaks

rtracklayer::export.bed(row.names(unchanging_reg), con = "unchanging_ATAC_peaks.bed")

# up_annotated_peaks = annotatePeak(peak= "./up_H3K27me3_peaks.bed", tssRegion = c(-5000, 5000))
# 
# down_annotated_peaks = annotatePeak(peak= "./down_H3K27me3_peaks.bed", tssRegion = c(-5000, 5000))

# this will be the plotted up peaks
# plotAnnoPie(down_annotated_peaks)
# plotAnnoPie(up_annotated_peaks)


# now to read in the peak files
peak_files = c("./up_ATAC_peaks.bed", "./down_ATAC_peaks.bed", "./unchanging_ATAC_peaks.bed")

up_peak_file = readPeakFile(peakfile = "./up_ATAC_peaks.bed")

unchanging_peak_file = readPeakFile(peakfile = "./unchanging_ATAC_peaks.bed")
unchanging_annotated_peaks = annotatePeak(peak= unchanging_peak_file, tssRegion = c(-5000, 5000), TxDb = txdb)

up_annotated_peaks = annotatePeak(peak= up_peak_file, tssRegion = c(-5000, 5000), TxDb = txdb)
plotAnnoPie(up_annotated_peaks)

down_peak_file = readPeakFile(peakfile = "./down_ATAC_peaks.bed")
down_annotated_peaks = annotatePeak(peak= down_peak_file, tssRegion = c(-5000, 5000), TxDb = txdb)

plotAnnoPie(down_annotated_peaks)



plotDistToTSS(down_annotated_peaks,
              title="Distribution of ATAC peaks\nrelative to TSS")


# diff_peaks = GRanges()
# 
# # so i appended the two peak files to a granges list
# diff_peaks = append(diff_peaks, up_peak_file)
# diff_peaks = append(diff_peaks, down_peak_file)
# 
# reduce(diff_peaks)

# peak_files = c("./up_H3K27me3_peaks.bed", "./down_H3K27me3_peaks.bed", "./unchanging_H3K27me3_peaks.bed")
# peak_files = c(up_peak_file = "./up_ATAC_peaks.bed", down_peak_file = "./down_ATAC_peaks.bed", unchanging_peak_file = "./unchanging_ATAC_peaks.bed")
```


```{r}
# ABOVE IS USING THE WRONG PEAKS NOT FROM THE FINAL PIPELINE WORKFLOW
# i was using the wrong peak files above
peak_files = c(up_peak_file = "./nextflow_R_script_outputs/ATAC/up_ATAC_dH1scr_regulated_peaks.bed", down_peak_file = "./nextflow_R_script_outputs/ATAC/down_ATAC_dH1scr_regulated_peaks.bed", unchanging_peak_file = "./nextflow_R_script_outputs/ATAC/unchanging_ATAC_dH1scr_regulated_peaks.bed")
# this is using the hg38 as txdb
annotated_peaks_list = lapply(peak_files, annotatePeak, tssRegion = c(-5000, 5000), TxDb = txdb) 


# now get the number of total annotated peaks
# down_annotated_df <- as.data.frame(down_annotated_peaks)
down_annotated_df <- as.data.frame(annotated_peaks_list$down_peak_file)
#num_down_annotated_peaks = count(down_annotated_df)[[1]]
num_down_annotated_peaks = length(down_annotated_df[,1])


# up_annotated_df <- as.data.frame(up_annotated_peaks)
up_annotated_df <- as.data.frame(annotated_peaks_list$up_peak_file)
num_up_annotated_peaks = count(up_annotated_df)[[1]]

# unchanging_annotated_df <- as.data.frame(unchanging_annotated_peaks)
unchanging_annotated_df <- as.data.frame(annotated_peaks_list$unchanging_peak_file)
num_unchanging_annotated_peaks = count(unchanging_annotated_df)[[1]]

# then plot with bar because it uses a ggplot object
plotAnnoBar(annotated_peaks_list) + ggtitle("Percentage of ATAC peaks in Genomic Regions", subtitle = paste("Number of down peaks annotated ", num_down_annotated_peaks, "\nNumber of up peaks annotated ",num_up_annotated_peaks, "\nNumber of unchanging peaks annotated ", num_unchanging_annotated_peaks ) )

# ggsave("annotation_bar_graph_ATAC_peaks.png", plot = last_plot())

vennpie(down_annotated_peaks)
upsetplot(down_annotated_peaks, vennpie=TRUE)

```



```{r}
# summary(annotated_peaks_list)
# 
# up_annotated_df$annotation == "Distal Intergenic" 
# 
# grepl("^Exon", up_annotated_df$annotation) # intron and exon will need to use grepl
# 
# up_annotated_df$annotation
# 
# # this looks through the annotation column in the dataframe and compares it to see which match the unique names found in the list of features included in the annotation
# # I can probably split string at the first space to get only the important unique features
# up_annotated_df$annotation == unique(up_annotated_df$annotation)[2]
# 
# 
# unique(gsub("\\s*\\(.*\\)", "",up_annotated_df$annotation))

getwd()
# making a for loop that will loop through the list of features 
# these are the up peaks, then will do the down peaks

unique_up_features = unique(gsub("\\s*\\(.*\\)", "",up_annotated_df$annotation))

for (i in c(1:length(unique_up_features))) {
  
  new_df = up_annotated_df[grepl(unique_up_features[i], up_annotated_df$annotation),]
  
  #new_up_annotated_df[grepl(paste0("^", strsplit(unique(up_annotated_df$annotation), ' ')[[i]][1]), up_annotated_df$annotation)]
    
  # replacing space with underscore
  no_apostrophy = gsub("'", " ", unique_up_features[i])
  
  feature_name_no_space = gsub(" ", "_", no_apostrophy)
  
  fileName_out = paste0("./up_ATAC_annotated_peaks_", feature_name_no_space, "_feature.bed")
  print(fileName_out)
  
  # rtracklayer::export.bed(new_df, con = fileName_out)
  
  write.table(new_df, file = fileName_out, sep = '\t', quote = FALSE, row.names = FALSE, col.names = FALSE )
  
  # export.bed below does not give the correct width of peaks
  
  # so I can use rtrack layer export properly, I need to make a granges object from the df
  # new_gr = makeGRangesFromDataFrame(new_df)
  
  # extra columns to include 
  # it seems like this doesnt include extraCols in export.bed
  # other_cols = c("annotation", "geneChr", "geneStart", "geneEnd", "geneLength", "geneStrand", "geneId", "transcriptId", "distanceToTSS")
  
  # rtracklayer::export.bed(new_gr, con =  fileName_out)
    
  
  
}


# now the for loop for the unchanging atac peaks

unique_up_features = unique(gsub("\\s*\\(.*\\)", "",unchanging_annotated_df$annotation))

for (i in c(1:length(unique_up_features))) {
  
  new_df = unchanging_annotated_df[grepl(unique_up_features[i], unchanging_annotated_df$annotation),]
  
  #new_up_annotated_df[grepl(paste0("^", strsplit(unique(up_annotated_df$annotation), ' ')[[i]][1]), up_annotated_df$annotation)]
    
  # replacing space with underscore
  no_apostrophy = gsub("'", " ", unique_up_features[i])
  
  feature_name_no_space = gsub(" ", "_", no_apostrophy)
  
  fileName_out = paste0("./unchanging_ATAC_annotated_peaks_", feature_name_no_space, "_feature.bed")
  print(fileName_out)
  
  # rtracklayer::export.bed(new_df, con = paste0("unchanging_ATAC_annotated_peaks_", feature_name_no_space, "_feature.bed"))
  
  write.table(new_df, file = fileName_out, sep = '\t', quote = FALSE, row.names = FALSE , col.names = FALSE)
  
  # export.bed below does not give the correct width of peaks
  
  # so I can use rtrack layer export properly, I need to make a granges object from the df
  # new_gr = makeGRangesFromDataFrame(new_df, keep.extra.columns = TRUE)
  
  # rtracklayer::export.bed(new_gr, con =  fileName_out)
  
  
}


# test

# unique_up_features[4]
# 
# length(which(grepl( paste0("^",unique_up_features[4]), up_annotated_df$annotation) == FALSE))
# 
# keep = which(grepl( paste0("^",unique_up_features[4]), up_annotated_df$annotation) == TRUE)
# 
# keep
# 
# test_df = up_annotated_df[keep,]
# 
# length(test_df[,1])
# 
# up_annotated_df$annotation
# 
# # strsplit(unique(up_annotated_df$annotation), ' ')[[i]][1]
# 
# grepl(paste0("^", strsplit(unique(up_annotated_df$annotation), ' ')[[2]][1]), up_annotated_df$annotation)
# 
# another_test_df = up_annotated_df[paste0("^",unique_up_features[4]), up_annotated_df$annotation),]
# 
# gsub(" ", "_", unique_up_features[2])

```

# now below I can calculate the total genomic range covered by the atac peaks around all the different features
```{r}

# make a pie chart where I show the total region covered by each feature 

# I can make a for loop and store the total genomic region of each peak feature into a dataframe outside of the loop
# then make the pie chart

sep_up_ann_atac_features_files = list.files(path = "./", pattern = "up.*ATAC_annotated.*feature.bed")
sep_up_ann_atac_features_files 
# getwd()

sep_unchanging_ann_atac_features_files = list.files(path = "./", pattern = "unchanging.*ATAC_annotated.*feature.bed")
sep_unchanging_ann_atac_features_files


genomic_total_features_df = DataFrame()

for (i in c(1:length(sep_up_ann_atac_features_files ))) {
  
  # I should grab the token that represents the feature
  file_name = sep_up_ann_atac_features_files[i]
  
  name_tokens = unlist(str_split(file_name, pattern = "peaks_"))
  
  final_name_feature = gsub(x = name_tokens[2], pattern = "_feature.bed" , replacement = "")
  
  peak_type = unlist(str_split(name_tokens[1], pattern = "_"))
  print(peak_type)
  # then I should use granges to get the width, but if the dataframe already has the width info in a column just take that and sum all the rows in that column.
  
  # rtracklayer::import.bed(file_name)
  feature_table = read.table(file_name, row.names = NULL)
  
  feature_genomic_range = sum(feature_table$V4)
  
  # then save to a dataframe outside the for loop
  genomic_total_features_df[final_name_feature] = feature_genomic_range
  # genomic_total_features_df["peak_type"] = peak_type[1]
  
}


unchanging_genomic_total_features_df = DataFrame()
# now unchanging
for (i in c(1:length(sep_unchanging_ann_atac_features_files ))) {
  
  # I should grab the token that represents the feature
  file_name = sep_unchanging_ann_atac_features_files[i]
  
  name_tokens = unlist(str_split(file_name, pattern = "peaks_"))
  
  final_name_feature = gsub(x = name_tokens[2], pattern = "_feature.bed" , replacement = "")
  
  peak_type = unlist(str_split(name_tokens[1], pattern = "_"))
  print(peak_type)
  # then I should use granges to get the width, but if the dataframe already has the width info in a column just take that and sum all the rows in that column.
  
  # rtracklayer::import.bed(file_name)
  feature_table = read.table(file_name, row.names = NULL)
  
  feature_genomic_range = sum(feature_table$V4)
  
  # then save to a dataframe outside the for loop
  unchanging_genomic_total_features_df[final_name_feature] = feature_genomic_range
  # genomic_total_features_df["peak_type"] = peak_type[1]
  
}



up_final_df = data.frame(t(genomic_total_features_df))
up_final_df["peak_type"] = "up"

unchanging_final_df = data.frame(t(unchanging_genomic_total_features_df))
unchanging_final_df["peak_type"] = "unchanging"

# now to merge the two dataframes
combined_up_unchanging_atac_df = rbind(up_final_df, unchanging_final_df)

# get the geometric mean of the values column and then divide each by that geometric mean
geometric_mean_of_values = exp(mean(log(combined_up_unchanging_atac_df$value)))


# vector divide
combined_up_unchanging_atac_df$value = combined_up_unchanging_atac_df$value/geometric_mean_of_values

# or we can try percentage
# total_of_values = sum(combined_up_unchanging_atac_df$value)

# combined_up_unchanging_atac_df$value = (combined_up_unchanging_atac_df$value/total_of_values)*100

ggplot(combined_up_unchanging_atac_df, aes(x= value, y = peak_type, fill = group_name, colour =))+
  geom_bar(stat="identity", position = "fill", width=1) +
  ggtitle("genomic coverage of ATAC-seq peaks")+
  scale_fill_brewer(palette = "Set2")+
  theme_minimal()
  # coord_polar("y", start=0)

ggsave(filename = "atac_peak_annotation_by_genomic_coverage.pdf", plot = last_plot(), device = "pdf" )
# for (x in sep_up_ann_atac_features_files) {
#   print(x)
# }

```

