---
title: "R Notebook"
output: html_notebook
---


# first let's load the necessary files
```{r}

library(rtracklayer)
library(GenomicRanges)

# the two c-score condition files

# h1low c-score regions
h1low_corrected_cscore = rtracklayer::import.bed("/lustre/fs4/risc_lab/scratch/iduba/linker-histone/HiC/KH3/deepseq/combinedreps/chroms/low-corrected-cscore.bedgraph")

scrm_corrected_cscore = rtracklayer::import.bed("/lustre/fs4/risc_lab/scratch/iduba/linker-histone/HiC/KH3/deepseq/combinedreps/chroms/scr-corrected-cscore.bedgraph")

```


# now lets load the proseq data
```{r}

# extraCols_gene_file = c(transcript = "string", gene_name = "string", strand = " string", gene_id = "string")

# extraCols_gene_file = c(gene_id = "character")

up_gene_regions_and_name = rtracklayer::import.bed("/lustre/fs4/home/rjohnson/pipelines/peak_calling_analysis_pipeline/proseq_data/proseq_up_genesID_with_coord_no_stats.bed" )

unchanging_gene_regions_and_name = rtracklayer::import.bed("/lustre/fs4/home/rjohnson/pipelines/peak_calling_analysis_pipeline/proseq_data/proseq_unchanging_genesID_with_coord_no_stats.bed" )

```

# now I need to save the new genes that contain only the tss
```{r}

up_gene_regions_and_name_TSS_only = GenomicRanges::resize(up_gene_regions_and_name, width = 1)

# now the unchanging genes 
unchanging_gene_regions_and_name_TSS_only = GenomicRanges::resize(unchanging_gene_regions_and_name, width = 1)



```

# now to find which compartment scores intersect the gene tss
# I need to keep the info on which compartment the gene was found in and record that compartment score
```{r}

# looks like its still best to use find overlaps

h1low_up_gene_overlap_hits = findOverlaps(h1low_corrected_cscore, up_gene_regions_and_name_TSS_only )

# the next two lines of code gets the index of which subject was found in the index of which query
# then I need to get the up gene tss that actually fell in a compartment and save by indexing by the hits of the subject
up_genes_in_h1low_compartments = up_gene_regions_and_name_TSS_only[subjectHits(h1low_up_gene_overlap_hits)]

# indexing by the hits of the query
h1low_compartments_that_had_up_genes = h1low_corrected_cscore[queryHits(h1low_up_gene_overlap_hits)]

# now take the meta data col that contains the compartment score and put it on the granges object that has the gene tss and gene id/name
h1low_c_score_for_genes = mcols(h1low_compartments_that_had_up_genes)

up_genes_in_h1low_compartments@elementMetadata["h1low_c-score"] = h1low_c_score_for_genes

# this should be the gene tss, with the gene name/id, with the compartment score for h1low
up_genes_in_h1low_compartments

# now I need to find the up genes that intersect the scrambled compartments which might be a different set of genes all together.
# so that means I would need to do an intersection of the up gens in h1low compartments with the up genes in the scrm compartments

# once I have that, I should be able to get the delta c-score and make a violin plot 

```

# now I need to find the up genes that intersect the scrambled compartments which might be a different set of genes all together.
# so that means I would need to do an intersection of the up gens in h1low compartments with the up genes in the scrm compartments

# once I have that, I should be able to get the delta c-score and make a violin plot 
```{r}



scrm_up_gene_overlap_hits = findOverlaps(scrm_corrected_cscore, up_gene_regions_and_name_TSS_only )

# indexing subject hits
up_genes_in_scrm_compartments = up_gene_regions_and_name_TSS_only[subjectHits(scrm_up_gene_overlap_hits)]

# indexing by the hits of the query
scrm_compartments_that_had_up_genes = scrm_corrected_cscore[queryHits(scrm_up_gene_overlap_hits)]

# now take the meta data col that contains the compartment score and put it on the granges object that has the gene tss and gene id/name
scrm_c_score_for_genes = mcols(scrm_compartments_that_had_up_genes)

up_genes_in_scrm_compartments@elementMetadata["scrm_c-score"] = scrm_c_score_for_genes

# this should be the gene tss, with the gene name/id, with the compartment score for h1low
up_genes_in_scrm_compartments


```


# it looks like we have the same number of up genes found in each of the conditions
# this makes it easy to just put the columns side by side and make a new column with the delta of h1low minus scrm
```{r}

# just make a new object that copies the h1low up genes with c-score

final_up_genes_with_h1low_scrm_cscores = up_genes_in_h1low_compartments

final_up_genes_with_h1low_scrm_cscores

# now just take the scrm cscore column and put it in this final object
final_up_genes_with_h1low_scrm_cscores@elementMetadata["scrm_c-score"] = mcols(up_genes_in_scrm_compartments)[2]

final_up_genes_with_h1low_scrm_cscores

# now do the subtraction of h1low and scrm c-scores

final_up_genes_with_h1low_scrm_cscores@elementMetadata["delta_h1low_scrm_cscore"] = as.numeric(final_up_genes_with_h1low_scrm_cscores@elementMetadata$`h1low_c-score`) - as.numeric(final_up_genes_with_h1low_scrm_cscores@elementMetadata$`scrm_c-score`)

# so now this is the final output for the up proseq genes
final_up_genes_with_h1low_scrm_cscores

```


###################################
# this section will be the unchanging genes

# now to find which compartment scores intersect the gene tss
# I need to keep the info on which compartment the gene was found in and record that compartment score
```{r}

# looks like its still best to use find overlaps

h1low_unchanging_gene_overlap_hits = findOverlaps(h1low_corrected_cscore, unchanging_gene_regions_and_name_TSS_only )

# the next two lines of code gets the index of which subject was found in the index of which query
# then I need to get the up gene tss that actually fell in a compartment and save by indexing by the hits of the subject
unchanging_genes_in_h1low_compartments = unchanging_gene_regions_and_name_TSS_only[subjectHits(h1low_unchanging_gene_overlap_hits)]

# indexing by the hits of the query
h1low_compartments_that_had_unchanging_genes = h1low_corrected_cscore[queryHits(h1low_unchanging_gene_overlap_hits)]

# now take the meta data col that contains the compartment score and put it on the granges object that has the gene tss and gene id/name
h1low_c_score_for_genes_unchanging = mcols(h1low_compartments_that_had_unchanging_genes)

unchanging_genes_in_h1low_compartments@elementMetadata["h1low_c-score_unchanging"] = h1low_c_score_for_genes_unchanging

# this should be the gene tss, with the gene name/id, with the compartment score for h1low
unchanging_genes_in_h1low_compartments

# now I need to find the up genes that intersect the scrambled compartments which might be a different set of genes all together.
# so that means I would need to do an intersection of the up gens in h1low compartments with the up genes in the scrm compartments

# once I have that, I should be able to get the delta c-score and make a violin plot 

```

# now I need to find the up genes that intersect the scrambled compartments which might be a different set of genes all together.
# so that means I would need to do an intersection of the up gens in h1low compartments with the up genes in the scrm compartments

# once I have that, I should be able to get the delta c-score and make a violin plot 
```{r}



scrm_unchanging_gene_overlap_hits = findOverlaps(scrm_corrected_cscore, unchanging_gene_regions_and_name_TSS_only )

# indexing subject hits
unchanging_genes_in_scrm_compartments = unchanging_gene_regions_and_name_TSS_only[subjectHits(scrm_unchanging_gene_overlap_hits)]

# indexing by the hits of the query
scrm_compartments_that_had_unchanging_genes = scrm_corrected_cscore[queryHits(scrm_unchanging_gene_overlap_hits)]

# now take the meta data col that contains the compartment score and put it on the granges object that has the gene tss and gene id/name
scrm_c_score_for_genes_unchanging = mcols(scrm_compartments_that_had_unchanging_genes)

unchanging_genes_in_scrm_compartments@elementMetadata["scrm_c-score_unchanging"] = scrm_c_score_for_genes_unchanging

# this should be the gene tss, with the gene name/id, with the compartment score for h1low
unchanging_genes_in_scrm_compartments


```


# it looks like we have the same number of up genes found in each of the conditions
# this makes it easy to just put the columns side by side and make a new column with the delta of h1low minus scrm
```{r}

# just make a new object that copies the h1low up genes with c-score

final_unchanging_genes_with_h1low_scrm_cscores = unchanging_genes_in_h1low_compartments

final_unchanging_genes_with_h1low_scrm_cscores

# now just take the scrm cscore column and put it in this final object
final_unchanging_genes_with_h1low_scrm_cscores@elementMetadata["scrm_c-score_unchanging"] = mcols(unchanging_genes_in_scrm_compartments)[2]

final_unchanging_genes_with_h1low_scrm_cscores

# now do the subtraction of h1low and scrm c-scores

final_unchanging_genes_with_h1low_scrm_cscores@elementMetadata["delta_h1low_scrm_cscore_unchanging"] = as.numeric(final_unchanging_genes_with_h1low_scrm_cscores@elementMetadata$`h1low_c-score_unchanging`) - as.numeric(final_unchanging_genes_with_h1low_scrm_cscores@elementMetadata$`scrm_c-score_unchanging`)

# so now this is the final output for the up proseq genes
final_unchanging_genes_with_h1low_scrm_cscores

```



# now lets try the plotting
```{r}

# the data needs to be a dataframe

final_up_genes_cscore_df = GenomicRanges::as.data.frame(final_up_genes_with_h1low_scrm_cscores)
final_up_genes_cscore_df

final_unchanging_cscore_df = GenomicRanges::as.data.frame(final_unchanging_genes_with_h1low_scrm_cscores)
final_unchanging_cscore_df

just_the_cscores_df = data.frame(final_up_genes_cscore_df$delta_h1low_scrm_cscore, final_unchanging_cscore_df$delta_h1low_scrm_cscore_unchanging, fix.empty.names = TRUE)
just_the_cscores_df

# here is the wilcox test (mann whitney) to see if the two observations are significant or not
wilcox_results_unchanging_up_delta_cscore = wilcox.test(final_up_genes_cscore_df$delta_h1low_scrm_cscore, final_unchanging_cscore_df$delta_h1low_scrm_cscore_unchanging, correct = TRUE, conf.int = TRUE, paired = FALSE)

pval_between_up_and_unchanging_tss_cscore = wilcox_results_unchanging_up_delta_cscore$p.value
pval_between_up_and_unchanging_tss_cscore

library(ggplot2)
library(ggpubr)

up_df = final_up_genes_cscore_df[c("name", "delta_h1low_scrm_cscore")]
up_df["group"] = "up_genes"
unchanging_df = final_unchanging_cscore_df[c("name", "delta_h1low_scrm_cscore_unchanging")]
library(dplyr)
unchanging_df = rename(unchanging_df, delta_h1low_scrm_cscore = delta_h1low_scrm_cscore_unchanging) 
unchanging_df["group"] = "unchanging_genes"

# now to make one df
final_combined_df = rbind(up_df, unchanging_df)


ggplot() +
  ggviolin( data = final_combined_df, x = "group", y = "delta_h1low_scrm_cscore" ) +
  scale_y_continuous(limits = c(-2,2), n.breaks = 20)+
  stat_compare_means( method = "wilcox.test")


delta_cscore_violin_plot_up_unchanging_proseq_genes+
  stat_compare_means( method = "wilcox.test", comparisons = list(c( "unchanging_genes", "up_genes")))
# now that I have the plot for the up genes delta c-score, I just have to ge the unchanging genes
# delta_cscore_violin_plot_up_unchanging_proseq_genes = ggplot() +
#   geom_violin( data = final_up_genes_cscore_df, aes(x = "up_genes", y = delta_h1low_scrm_cscore) ) +
#   geom_violin(data = final_unchanging_cscore_df, aes(x = "unchanging_genes", y = delta_h1low_scrm_cscore_unchanging ) ) +
#   scale_y_continuous(limits = c(-2,2), n.breaks = 20)+
#   stat_compare_means(paired = TRUE, method = wilcox.test(), )

delta_cscore_violin_plot_up_unchanging_proseq_genes

ggsave(filename = "delta_cscore_violin_plot_up_unchanging_genes.png", delta_cscore_violin_plot_up_unchanging_proseq_genes)

# ggplot(final_unchanging_cscore_df, aes(x = "unchanging_genes", y = delta_h1low_scrm_cscore_unchanging ))+
#   geom_violin(scale = "area")+
#   geom_violin(data = final_up_genes_cscore_df, aes(x = "up_genes", y = delta_h1low_scrm_cscore))
#   scale_y_continuous(limits = c(-2,2), n.breaks = 20)

```




# scrap paper below

# lets see the man whitney test to see significance
```{r}

test_results = wilcox.test(final_up_genes_cscore_df$delta_h1low_scrm_cscore, final_unchanging_cscore_df$delta_h1low_scrm_cscore_unchanging, correct = TRUE, conf.int = TRUE, paired = FALSE)

test_results$p.value


ggplot() +
  geom_violin( data = final_up_genes_cscore_df, aes(x = "up_genes", y = delta_h1low_scrm_cscore) ) +
  geom_violin(data = final_unchanging_cscore_df, aes(x = "unchanging_genes", y = delta_h1low_scrm_cscore_unchanging ) ) +
  stat_compare_means(paired = TRUE, method = "wilcox.test", label = "p.signif") +
  scale_y_continuous(limits = c(-2,2), n.breaks = 20)

```



# play around with understanding how to get just the chromosome and start interval
```{r}

up_gene_regions_and_name

GenomicRanges::resize(up_gene_regions_and_name, width = 1)


```

# checking the resize
```{r}

# lets go with the width of 1 to get the tss
GenomicRanges::resize(up_gene_regions_and_name, width = 1)

```

